Paper

### 2018-02-20 Paper.Rmd
Updating the pipeline to make it more generic

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')

data = '~/Data/Ellie'
tile.w=c(5e06, 'arms')
#tile.w=5e+06
#tile.w=1e7
#tile.w='arms'

includeDemo = F

data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)

load(grep('All_patients.Rdata', analysis.files, value=T), verbose=T)

## Hospital.Research.ID info file
patient.file = grep('All_patient_info.xlsx', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('Demographics_full.xlsx', data.files, value=T)

@@

validation.patient.data = patient.data[sum.validation.info$Hospital.Research.ID]
length(validation.patient.data)

patient.data = patient.data[sum.patient.data$Hospital.Research.ID]
length(patient.data)

patient.data = patient.data[!is.na(names(patient.data))]
names(patient.data)

sum.patient.data = as.data.frame(subset(sum.patient.data, Hospital.Research.ID %in% names(patient.data))) ## one patient was removed from the study
nrow(sum.patient.data)

@@

# Segment size: `r as.character(tile.w)` with demographics: `r includeDemo`

@@

```{r, echo=F}
x = do.call(rbind.data.frame, lapply(patient.data, function(df) df$info[,c('Status','p53.Status')]))
grid.arrange(
  ggplot(patient.info, aes(p53.Status, fill=Status)) + geom_bar(position = position_dodge()) + labs(title='p53 Staining') + plot.theme,
  tableGrob( 
    rbind( 'Non-progressor'=summary(subset(x, Status == 'NP')$p53.Status), 
           'Progressor'=summary(subset(x, Status == 'P')$p53.Status))), 
  nrow=2, as.table=T, heights=c(4,1) )
```

@@

```{r labelsPNP, echo=F, message=F, include=T}
## binomial: dysplasia 1, BE 0
labels = unlist(sapply(patient.data, function(df) {
  df$info = plyr::arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

pts = do.call(rbind, lapply(patient.data, function(df) {
  df$info = plyr::arrange(df$info, Endoscopy.Year, Pathology)
  cbind(df$info[,c('Hospital.Research.ID','Samplename')])
}))
rownames(pts) = 1:nrow(pts)

cache.dir = paste(data, 'Analysis', sub('\\+','',paste(tile.w, collapse='_')), sep='/')

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
rm(plots, coefs, performance.at.1se, models, cv.patient)

@@

n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))
# Average cn value
globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
globalCNdf$mean.value = globalCNdf$value # but not going to use it

@@

## Dysplasia only
dys.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology %in% c('LGD','HGD','IMC'))$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology %in% c('LGD','HGD','IMC'))$variable)),')', sep=''))

#globalnHGDCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology %in% c('LGD','HGD','IMC')), labeller(Status=dys.status, chr=label_value)) + labs(title='LGD/HGD/IMC Only')

## No HGD
# `%nin%` = Negate(`%in%`)
# 
# nhgd.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology %nin% c('HGD','IMC'))$variable)), ')', sep=''),
#               'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology %nin% c('HGD','IMC'))$variable)),')', sep=''))
# globalnHGDCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology %nin% c('HGD','IMC')), labeller(Status=nhgd.status, chr=label_value)) + labs(title='Excluding HGD/IMC')

@@

#globalnHGDCN.plot + theme(legend.position = 'none')

@@

# dfpts = subset(globalCNdf, Patient %in% c('57','76') & seg.type != 'arms')
# #dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
# 
# pt.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
#               'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
# genomePtCN.plot = cn.mtn.plot(dfpts, labeller(Status = pt.status))
# 
# dfpts = subset(globalCNdf, Patient %in% c('57','76') & Pathology == 'BE' & seg.type != 'arms')
# #dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
# 
# ptbe.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
#               'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
# genomePtBECN.plot = cn.mtn.plot(dfpts,  labeller(Status = ptbe.status )) + labs(title='NDBE samples only')
# 
# genomePtCN.plot + theme(legend.position = 'none')
# genomePtBECN.plot + theme(legend.position = 'none')

#cn.mtn.plot(dfpts,  labeller(Status = ptbe.status ), type='bar') + labs(title='NDBE samples only')

#a = sample( subset(subset(patient.info, Patient == 57),  Endoscopy.Year == min(Endoscopy.Year) )$Samplename, 1)
#b = sample( subset(subset(patient.info, Patient == 76), Endoscopy.Year == min(Endoscopy.Year) & Pathology %nin% c('HGD','IMC'))$Samplename, 1)

#ind.status <- c('Non-Progressor'='Patient 57 (n=1)','Progressor'='Patient 76 (n=1), excl. HGD/IMC')
#cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(Status = ind.status), type='bar') + labs(title="Individual samples")
```

@@

# cx was generated on raw seg values
plot(dysplasia.df[names(labels) ,'cx'], col=labels+1)

# remove old cx...see what new one would look like
dysplasia.df = dysplasia.df[,-grep('cx',colnames(dysplasia.df))]

# general variance per sample after the samples have been mean normalized, ignores segment length
cx = apply(dysplasia.df, 1, function(x) var(x))

@@

tt = t.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
ggplot(x, aes(Status, cx, fill=Status)) + 
  geom_jitter(width=0.2, color='grey39') + 
  geom_violin(outlier.colour = NA, alpha=0.8) +
  labs(x='', y='Variance', title='Per sample variance', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
  plot.theme

rocX = pROC::roc(Status ~ cx, data=x, auc=T, ci=T, of='thresholds')
roc.plot(rocX, title='Per sample variance') + plot.theme

@@

@@

```{r echo=F, fig.height=5, fig.width=5}
library(ggfortify)
autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme
```

@@

```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}

classification.rate = as_tibble(matrix(ncol=5, nrow=0, dimnames=list(c(), c('model','mean','sme','stable.coef', 'all.coef'))))

mp = model.performance(performance.at.1se, coefs, lims=c(0.7,0.8))
mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))

@@

nohgd.coefs = coefs[[select.alpha]]
#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
```

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs, lims=c(0.69,0.8))
mp$plot + labs(title='No HGD/IMC samples')

@@

```{r loonohgd, echo=F, message=F, warning=T}

file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=F)
} else {
  stop(paste("No file", file, "found"))
}
```

```{r, echo=F, message=F, fig.width=5, fig.height=5}

  pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)

@@

  ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')

@@

@@

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs, lims=c(0.69,0.8))
mp$plot + labs(title='No LGD/HGD/IMC samples')

@@

ggplot(classification.rate, aes(model, mean, group=model, color=model)) + 
  geom_text( aes(label=round(mean, 3)), hjust=-0.5, size=5, color='gray39', show.legend = F) +
  geom_text( aes(y=mean+sme*2, label=paste(all.coef, "\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
  geom_point(size=4) +
  labs(x='', y='Classificaiton rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') + 
  ylim(0.7,0.8)

@@

qq = quantile(ch$hazards, seq(0,1,.15))
ggplot(ch, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') + plot.theme

@@

write.table(ch, quote=F, sep='\t', row.names = F, col.names = T, file=paste( paste(tile.w, collapse='_')  ,'_features.txt'))
```

@@

globalCN.plot + geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
  geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

globalBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

# ch$ymax = ifelse(ch$coef > 0, max(genomePtCN.plot$data$value), min(genomePtCN.plot$data$value))
# chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))
# genomePtCN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

# genomePtBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

@@

```{r echo=F, fig.height=5, fig.width=6}
globalCNdf$g.start = globalCNdf$start
globalCNdf$g.end = globalCNdf$end
for (i in 2:nrow(chr.lengths)) {
  rows = which(globalCNdf$chr == chr.lengths[i, 'chrom'])
  globalCNdf[rows, 'g.start'] = globalCNdf[rows, 'start'] + chr.lengths[i-1, 'genome.length']
  globalCNdf[rows, 'g.end'] = globalCNdf[rows, 'end'] + chr.lengths[i-1, 'genome.length']
}

globalCNdf = arrange(globalCNdf, chr, start)
globalCNdf$segment = with(globalCNdf, paste(chr, ':', start, '-', end, sep=''))

globalCNdf$chr = factor(globalCNdf$chr, levels=c(1:22), ordered = T)
globalCNdf$Status = factor(globalCNdf$Status)

p1 = ggplot(subset(globalCNdf, seg.type != 'arms'), aes(segment, variable, fill=value, group=Status)) + geom_tile() +
  scale_fill_gradient2(high='red3', low='green4', mid='white')  + 
  facet_grid(~chr, scales = 'free_x', space = 'free') + plot.theme + labs(x='',y='Samples') + 
  theme(axis.ticks=element_blank(), axis.text=element_blank(), panel.spacing.x = unit(0, 'lines'), legend.position='bottom')

p1leg = get.legend(p1)
p1 = p1 + theme(legend.position = 'none')

p2 = ggplot(subset(globalCNdf, seg.type != 'arms'), aes(y=variable, group=Status)) + 
  geom_tile(aes(x=0, fill=Status), show.legend = F) + 
  labs(x=' ', y='') + scale_fill_manual(values=c('darkblue','darkred')) + scale_x_continuous(expand=c(0,0)) +
  plot.theme + theme( axis.line=element_blank(), axis.text=element_blank(), axis.ticks=element_blank(), panel.grid=element_blank())

grid.arrange(p1, p2, p1leg, ncol=2, nrow=2, widths=c(20,1), heights=c(10,1))

```{r echo=F, fig.height=4, fig.width=8}
ggplot(chh, aes(x=label,y=coef,fill=gl)) + geom_bar(stat='identity') + labs(x='Effects', y='') +
  coord_flip() + scale_fill_manual(values=c('darkgreen','darkred')) +
  plot.theme + theme(axis.line.x=element_blank(), axis.ticks=element_blank(),axis.text=element_blank(), panel.grid=element_blank(),panel.grid.major=element_blank(), legend.position='none')
```

@@

df = as.data.frame(performance.at.1se)
ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
  plot.theme

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'discovery'
roc$is = 'discovery'

droc = multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
ggsave(filename='plots/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)

@@

p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
    geom_line(data=subset(df, model == 'discovery'), color='red', size=1) +
    scale_x_reverse() +
    scale_color_brewer(palette='Blues') +
    labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
p

ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

ir = ggplot(insilico.rocs, aes(is, auc, color=split)) + ylim(0,1) +
  geom_hline(data=m, aes(yintercept=auc, linetype=line), color='grey32', show.legend=F) + 
  geom_point(size=2) + geom_errorbar(aes(ymin=auc.ci.min, ymax=auc.ci.max), width=0.3) + 
  geom_text(data=m, aes(label=paste(variable, round(auc, 2), sep='\n')), color='grey32', nudge_y=0, show.legend=F) + 
  scale_linetype_manual(values=c('dotdash','longdash')) +
  plot.theme + scale_color_brewer(palette='Set1', name='Type') +
  theme(axis.text.x=element_blank(), legend.position = 'bottom') +
  labs(y='AUC', x='Data splits', title='AUC per split', subtitle='20 insilico 20:69 splits')
ir
ggsave(filename='plots/auc_splits.png', plot=ir, width=7, height=8, units='in', dpi=300)

# To find the auc at a given spec or sens
#pROC::auc(roc, partial.auc=c(0.99,1), partial.acu.focus="sp", partial.auc.correct=T)

df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))

thresholds = 
  rbind.data.frame(
#    est=t(round(pROC::coords(roc, "best"),2)),
    max=t(round(pROC::coords(roc, 0.8),2)),
    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6),2)),
    'All samples'=t(round(pROC::coords(roc, 0.6),2))
)
thresholds$label=rownames(thresholds)

ggplot() + geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() + 
  geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
  geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
    paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(3,'Dark2')) + scale_color_manual(values=brewer.pal(3,'Dark2')) +
  labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme

@@

```{r pathroc, echo=F, fig.height=4, fig.width=4 }
#sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))

pathsamples = patient.info[,c('Pathology','Status')]

pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
pathroc$model = 'HGD & LGD'

hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
hgdpathroc$model = 'HGD Only'

@@

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoM, 'best')
#roc.plot(rocEndoM)
rocEndoM$model = 'Per Endoscopy Max'

rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoA, 'best')
#roc.plot(rocEndoA)
rocEndoA$model = 'Per Endoscopy Avg'

@@

@@

ggsave('plots/compare_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)

>>>>

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')
source('lib/data_func.R')

data = '~/Data/Ellie'
info.dir = '~/Data/Ellie/QDNAseq'
#tile.w=c(5e06, 'arms')
#tile.w=5e+06
#tile.w=1e7
#tile.w='arms'

#includeDemo = F

#data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
#analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)

## Hospital.Research.ID info file

patient.file = list.files(info.dir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
demo.file = list.files(info.dir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)

if (length(patient.file) != 1 | length(demo.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", info.dir))

@@

@@

# Segment size: 5e06 with arms

@@

```{r, echo=F}
grid.arrange(
  ggplot(patient.info, aes(p53.Status, fill=Status)) + geom_bar(position = position_dodge()) + labs(title='p53 Staining') + plot.theme,
  tableGrob( 
    rbind('Non-progressor'=summary(subset(patient.info, Status == 'NP')$p53.Status), 
           'Progressor'=summary(subset(patient.info, Status == 'P')$p53.Status))), 
  nrow=2, as.table=T, heights=c(4,1) )
```

@@

```{r labelsPNP, echo=F, message=F, include=T}
cache.dir = paste(data, 'Analysis/5e6_arms_disc', sep='/')

file = paste(cache.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))

load(file, verbose=F)
dim(dysplasia.df)
length(labels)

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=F)
rm(plots, coefs, performance.at.1se, models)

@@

n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))

# Average cn value
globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
globalCNdf$mean.value = globalCNdf$value # but not going to use it

@@

@@

@@

@@

cx = dysplasia.df[,'cx']

@@

tt = t.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
pv = ggplot(x, aes(Status, cx, fill=Status)) + 
  geom_jitter(width=0.2, color='grey39') + 
  geom_boxplot(outlier.colour = NA, alpha=0.8) +
  labs(x='', y='cx', title='Per sample cx scores', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
  plot.theme
pv

rocX = pROC::roc(Status ~ cx, data=x, auc=T, ci=T, of='thresholds')
roc.plot(rocX, title='Per sample cx') + plot.theme

ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

@@

ggsave('plots/eyebrows.png', plot=do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits')), width=10, height=10, units='in', dpi=300)

@@

```{r echo=F, fig.height=5, fig.width=5}
library(ggfortify)
autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme

ggsave('plots/glm_0.9.png', plot=autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme, width=5, height=5, units='in', dpi=300)
```

@@

```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}
classification.rate = as_tibble(matrix(ncol=5, nrow=0, dimnames=list(c(), c('model','mean','sme','stable.coef', 'all.coef'))))

mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
mp$plot

ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

@@

nohgd.coefs = coefs[[select.alpha]]
#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
```

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No HGD/IMC samples')
mp$plot

ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

@@

```{r loonohgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}
```

```{r, echo=F, message=F, fig.width=5, fig.height=5}
pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)

@@

p = ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
p

ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)

@@

#hist(hgd.samples$Prediction)

@@

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No LGD/HGD/IMC samples')
mp$plot

ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=30)

@@

p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
  geom_text( aes(label=round(mean, 2)), vjust=2, size=5, color='gray39', show.legend=F) +
  geom_text( aes(y=0.6, label=paste(all.coef, " coef\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
#  geom_point(size=2) +
  labs(x='', y='Classification rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') +
  ylim(0,1)
p

ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
```

@@

qq = quantile(ch$hazards, seq(0,1,.15))
p = ggplot(ch, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') + plot.theme 
p

ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)

@@

@@

p = globalCN.plot + geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
  geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')

ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

p = globalBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')

ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

@@

@@

names(performance.at.1se) = names(pg.samp)
df = as.data.frame(performance.at.1se)

p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
  plot.theme
p

ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'CN Model'
roc$is = 'discovery'

droc = multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
droc
ggsave(filename='plots/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)

@@

p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
    geom_line(data=subset(df, is == 'discovery'), color='red', size=1) +
    scale_x_reverse() +
    scale_color_brewer(palette='Blues') +
    labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
p
ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
thresholds = 
  rbind.data.frame(
#    est=t(round(pROC::coords(roc, "best"),2)),
    max=t(round(pROC::coords(roc, 0.8),2)),
    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6),2)),
    'All samples'=t(round(pROC::coords(roc, 0.6),2))
)
thresholds$label=rownames(thresholds)

# ggplot() + geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() + 
#   geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
#   geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
#     paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(3,'Dark2')) + scale_color_manual(values=brewer.pal(3,'Dark2')) +
#   labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme

@@

```{r pathroc, echo=F, fig.height=4, fig.width=8 }
#sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))

pathsamples = patient.info[,c('Pathology','Status')]
pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
pathroc$model = 'HGD & LGD'
ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
hgdpathroc$model = 'HGD Only'
ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoM, 'best')
rocEndoM$model = 'Per Endoscopy Max'
ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoA, 'best')
rocEndoA$model = 'Per Endoscopy Avg'
ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

roc$model = "CN Model"

@@

ggsave('plots/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)
